---
title: "My First Report"
author: "Sky Ng"
format:  
  html: 
    embed-resources: true
---

## Welcome to your practice report

This is a **Quarto** document. It mixes ordinary writing with live code, so your
report and your analysis always stay in sync. When you click **Render**, Quarto
runs the code below and builds a finished HTML page.

Try it now: change `author: "Your Name"` in the header above to your own name,
then click **Render**.

## A first look at the data

The code chunk below reads a small practice dataset from the `data/` folder and
shows the first few rows.

```{r}
patients <- read.csv("data/patients.csv")
head(patients)
```

## A first chart

```{r}
hist(patients$age,
     main = "Age distribution of practice patients",
     xlab = "Age (years)",
     col  = "steelblue")
```

## Your turn 

### Data Table Analysis
- The data table shows us the health and basic information of 6 humans
- Information Provided: id, age, sex, region, systolic_bp and diabetes

### Chart Analysis
- The chart represents the age distribution of practise patients
- Age (years) is the X axis heading, and frequency is the Y axis heading
- The age ranges with the highest frequency are: 40-50, 50-60 and 60-70
- the age range with the lowest frequency is 70-80

When you are happy with the result, **commit and sync** your work (the Week 2
course page walks you through this), then submit your repository link on Canvas.

## My Notes: What I have learnt
- syntax for markdown documents: such as how to **bold**, *italicize* and create lists
- how to embed a section that runs R code, and relating R functions (e.g summary())
- the process of rendering markdown files into HTML and PDFs, and how to download it

## One Small R Calculation
```{r}
result <- 5 + 10
print(result)
```