---
title: "Assignment 2 Starter Analysis"
format: html
---

## Setup

```{r}
#| message: false
#| warning: false
library(tidyverse)  # AI-EDIT(2026-06-23): tidyverse-default — consolidated core library() calls into library(tidyverse)
```

## Load Data

```{r}
nhanes <- read_csv("examples/nhanes-equity/data/nhanes_equity_v6.csv", show_col_types = FALSE)
```

## Inspect

```{r}
dim(nhanes)
names(nhanes)
glimpse(nhanes)
colSums(is.na(nhanes))
```

## AI Prompt Used

### Using the NHANES data frame `nhanes` and tidyverse functions, write two beginner-friendly R functions. The first should create an analysis-ready table for adults age 20 or older by filtering missing age, gender, and BMI values, assigning BMI categories, and summarizing participant counts and mean age by gender and BMI category. The second should accept a numeric vector and return its non-missing count, mean, and standard deviation in a tibble. Use the dataset's existing column names, demonstrate the first function with `nhanes` and the second with `nhanes$Age`, and briefly explain what each function returns.

## AI Audit Note
### See [ai-use-note.md](ai-use-note.md) for the AI-use audit trail and verification details.


## Analysis-ready table using at least three dplyr verbs.

### Description of the function: 
#### This function filters the NHANES data to adults with recorded age, gender, and BMI, then assigns each person to a BMI category. It returns a table showing the participant count and average age for each gender and BMI-category group.

#### code generated by AI, verified by running the function and checking the output for expected structure and values.

```{r}
# AI prompt: Create a dplyr function to categorize adult BMI and summarize by gender.
# Verified: Checked the NHANES column names and ran the function; it returned 8 groups.
bmi_categories_by_gender <- function(data) {
  data |>
    filter(!is.na(Age), Age >= 20, !is.na(Gender), !is.na(BMI)) |>
    select(Age, Gender, BMI) |>
    mutate(
      BMI_category = case_when(
        BMI < 18.5 ~ "Underweight",
        BMI < 25 ~ "Healthy weight",
        BMI < 30 ~ "Overweight",
        TRUE ~ "Obesity"
      )
    ) |>
    group_by(Gender, BMI_category) |>
    summarise(
      participant_count = n(),
      mean_age = mean(Age),
      .groups = "drop"
    )
}

bmi_summary <- bmi_categories_by_gender(nhanes)
bmi_summary
```


## Create a custom function (such as a function that returns mean and SD for a numeric variable)

### mean_sd() takes a numeric vector, ignores missing values, and returns its non-missing count, mean, and standard deviation; here, it is used to summarize NHANES participant ages

```{r}
# AI prompt: Create a reusable function that returns n, mean, and SD for a numeric vector.
# Verified: Ran it on nhanes$Age; it returned n = 105626, mean = 31.7, and SD = 24.9.
mean_sd <- function(x) {
  tibble(
    n = sum(!is.na(x)),
    mean = mean(x, na.rm = TRUE),
    sd = sd(x, na.rm = TRUE)
  )
}

mean_sd(nhanes$Age)
```


## Save Output

```{r}
# AI-EDIT: Save the analysis-ready BMI summary as the required CSV output.
dir.create("assignments/assignment02-r-with-ai/submission/output", recursive = TRUE, showWarnings = FALSE)
write_csv(bmi_summary, "assignments/assignment02-r-with-ai/submission/output/summary_table.csv")
```
